---
title: "Practice: Cause and community"
author: "Andrew Heiss"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    lib_dir: "libs"
    chakra: "libs/remark-latest.min.js"
    css: ["default", "css/ath-slides.css", "css/ath-jost-fonts.css", "css/animate.css"]
    seal: false
    anchor_sections: false
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
      navigation:
        scroll: false
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, 
                      fig.retina = 3, fig.align = "center")
```

```{r packages-data, include=FALSE}
library(countdown)
```

```{r xaringanExtra, echo=FALSE}
xaringanExtra::use_xaringan_extra(c("tile_view"))
```

class: center middle main-title section-title-2

# Practice day<br>.small[Cause and community]

.class-info[

**January 20, 2022**

.light[PMAP 3210: Introduction to Nonprofits<br>
Andrew Young School of Policy Studies
]

]

.corner-ribbon[**Do your check-in!**]

---

name: outline
class: title title-inv-5

# Plan for today

--

.box-8[.medium[Why do nonprofits even exist?]<br>.small[Theory part]]

--

.box-1.medium[How to make teamwork not suck]

--

.box-6.medium[Forming and norming]

--

.box-3.medium[Issue and cause selection]

---

layout: false
name: theory
class: center middle section-title section-title-8 animated fadeIn

# Why do<br>nonprofits even exist?<br>.smaller[(Theory part)]

---

layout: false
name: teams-not-suck
class: center middle section-title section-title-1 animated fadeIn

# How to make<br>teamwork not suck

---

layout: true
class: title title-1

---

# Teams are everywhere!

--

.box-inv-1[There's a whole world of scholars<br>who study what makes teams effective]


.box-1.small.sp-after[(Organizational behavior)]

--

.box-inv-1[Teams generally increase<br>productivity, quality, and morale]

.box-inv-1[They're a good thing!]

???

Teams are everywhere!

517 - Teams generally increase productivity, quality, and morale, but often managers see this success and order immediate adoption of work teams without any preparation—leads to failure, creates inefficient teams

---

# Few do them right

.box-inv-1.SMALL[Often, managers (and professors) see the success of teams<br>and make their employees (or students) adopt them<br>***without any preparation***]

--

.box-1.medium[This is bad!]

--

.box-inv-1[Trying to work in a team without any preparation<br>leads to failure and inefficiencies]

--

.box-inv-1[Teams can create public good deficiencies!]

---

layout: false
class: middle

.box-1.large[We're going to <br>do teams right]

---

layout: true
class: title title-1

---

# Stages of the team lifecycle

--

.box-inv-1.medium[Forming]

--

.box-inv-1.medium[Norming]

--

.box-inv-1.medium[Storming]

--

.box-inv-1.medium[Performing]

---

# Forming

--

.pull-left[
.box-inv-1.SMALL[Members aren't a team;<br>just individuals]

.box-inv-1.SMALL[Lots of uncertainty<br>about everything]

.box-inv-1.SMALL[Silence, self-consciousness, dependence, and superficiality reign supreme]
]

--

.pull-right[
.box-1.SMALL[Figure out the details<br>of the task or project]

.box-1.SMALL[Get questions answered]

.box-1.SMALL[Establish trust and<br>build relationships]

.box-1.SMALL[Establish clarity of purpose,<br>procedures, and expectations]
]

???

*Developing Management Skills: Chapter 9, Teams (part 1)* (from BYU MPA OB class)

At first team members aren't a team, but an aggregation of individuals sharing a common setting - there's uncertainty as to what is going to happen, what is expected, who everyone is, who will lead. No shared history, no unity

Silence, self consciousness, dependence, and superficiality reign supreme

In this stage, in order to overcome all that, a team must:

* Orient members and get questions answered
* Establish trust
* Establish relationships with the leaders
* Establish clarity of purpose, norms, procedures, and expectations

Not a time to rely on free and open discussion and consensus decision making - direction, clarity, and structure are needed

---

# Norming

--

.box-1.medium[Start building a team culture]

--

.box-inv-1.SMALL[Codify norms and expectations in a written charter]

--

.box-inv-1.SMALL[Start conforming to standards and expectations]

--

.box-inv-1.SMALL[Maintain unity and cohesion]

--

.box-inv-1.SMALL[Facilitate participation and empowerment]

--

.box-inv-1.SMALL[Show support to team members]

???

Create team culture, a sense of team - members start to work together, conforming to expectations of other team members, creating team character

522 - Team members work to fit in, support each other, get along

Members begin to value team opinion and goals over personal feelings of cooperativeness

Conformity to standards and expectations, ignoring disagreements

A team must:

* Maintain unity and cohesion
* Facilitate participation and empowerment
* Show support to team members
* Provide feedback on team and member performance

**But**, you **must** move on from this stage, otherwise groupthink sets in and drives out good decision making - no challenge to decisions - too much emphasis on preserving the team rather than making accurate, high quality decisions

---

# Storming

--

.pull-left[
.box-inv-1.SMALL[Harmony and success<br>can lead to dissension]

.box-inv-1.SMALL["All sunshine makes a desert"]

.box-inv-1.SMALL[Coalitions and cliques]

.box-inv-1.SMALL[**Free-riding**]
]

--

.pull-right[
.box-1.SMALL[Hold members to<br>expectations in the charter]

.box-1.SMALL[Manage conflict]

.box-1.SMALL[Legitimize productive<br>expressions of individuality]

.box-1.SMALL[Turn counterdependence<br>into interdependence]
]

???

Harmony and success can lead to dissension - "All sunshine makes a desert"

Coalitions and cliques form—there's competition between team members, disagreement with the leader—challenges to others' points of view

Main tasks:

* Manage conflict
* Legitimize productive expressions of individuality
* Turn counterdependence into interdependence
* Foster consensus building process

---

# Performing

--

.box-inv-1.sp-after[After forming, norming, and occasional storming,<br>the team can work at a high level of performance!]

--

.box-1.medium[Members show shared responsibility<br>and concern for one another]

???

Team has worked through each of the previous stages - is able to work at a high level of performance

Keep going at that level—how can we help members thrive? Foster continuous improvement and creativity? Build on core competence? Maintain high level of energy/

Members show shared responsibility and concern for one another, coach and assist one another

---

# Today's plan

--

.pull-left[
.box-1.less-medium[Forming]

.box-inv-1[Create a team name]

.box-inv-1[Get to know each other]

.box-inv-1[Get questions answered]
]

--

.pull-right[
.box-1.less-medium[Norming]

.box-inv-1[Establish roles<br>and expectations]

.box-inv-1[Write a charter]
]

---

# Today's goals

.box-inv-1.medium[Create a name]

.box-inv-1.medium[Create a charter]

.box-inv-1.medium[Settle on a community and issue and<br>preliminary nonprofit for final project]

???

Forming + norming

*Highly* recommend trying to do something together as a team that is wholly unrelated to the class and where you don't talk about the class whatsoever - bowling

Team name

---

layout: false
name: charter
class: center middle section-title section-title-6 animated fadeIn

# Forming<br>and norming

---

layout: false
name: issue-cause
class: center middle section-title section-title-3 animated fadeIn

# Issue and<br>cause selection
